---
title: "Markdown Basics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Some R Markdown syntax

Inserting line breaks into the document with either html tags or a slash followed by 2 spaces:

<br><br>

\   

<br>

Headers can be in different sizes:

# Super Big

## Big

### Average

#### Small

##### Tiny

###### I can't even see this.

\   

You can *italicize text* like this or  _like this_, you can **bold text** like that or __like this__. You can do **_combinations_** of these.  Sometimes you like to ~~strikethrough~~
text.  Or even write things in superscript^like this^.  

There are other syntaxy things you might want to write such as an endash -- , an emdash --- or an ellipsis ... . These are written using the following code `--`, `---`, `...`.


You can write plain text in this space and it appears as text in your final document.  If you want to start a new paragraph make sure you leave two spaces at the end of your line.  

This is a new paragraph.  



\   

You can write weblinks using the syntax: `[some text](weburl)` and it will look [like this ](https://www.youtube.com/watch?v=dQw4w9WgXcQ) in the text.


\ 

### Code in text

It's simple to write some code in plain text by putting the text inside tildes like this `head(df)`.  Sometimes you want to write chuncks of text in code. That can be done using three tildes either side of the code chunk - ensure that `echo=T`.

```{r, echo=T, eval=F}
head(df)
df$newvar <- c(1:5)
```

\  

\  

Text can also be written in a block quote.

> _No act of kindness, no matter how small, is ever wasted_ - Aesop.

\  

You may also want to write bullet lists or numbered lists and you can do that like this:

* unordered list
* item 2
 + sub-item 1
 + sub-item 2

1. ordered list
2. item 2
 + sub-item 1
 + sub-item 2 
 
 

 
 \  
 
Embed images

Embed videos


\  

You can add a linebreak with five dashes

-----

\  


### Playing around with HTML and CSS

\  

With a bit of practice you can add HTML and CSS syling elements to a Markdown document.  This is because you are essentially creating raw HTML from the Markdown language via R.  So, let's say you wanted to change the color of some text, <p style="color:red;"> let's do that like this by adding some inline CSS/HTML </p>.

\  


#### Math

It's possible to write inline equations by wrapping some math notation with `$` like this:  $A = \pi*r^{2}$  . That was written with the following code `$A = \pi*r^{2}$`. If you need guidance on how to write math notation - there are lots of [guides on the internet](http://csrgxtu.github.io/2015/03/20/Writing-Mathematic-Fomulars-in-Markdown/). [^1]


Other examples are Greek letters: `$\Sigma \; \alpha \; \gamma$`: $\Sigma \; \alpha \; \gamma$

Sums, products, and integrals: `$\sum_{i=1}^{\infty}x_i$` $\sum_{i=1}^{\infty}x_i$  

With outer limits: `$\prod\limits_{i=1}^{\infty}\frac{1}{x_i}$` $\prod\limits_{i=1}^{\infty}\frac{1}{x_i}$

Displayed as a block: `$$\int_{a}^{b}x^2dx$$` $$\int_{a}^{b}x^2dx$$


There are several ways of writing matrices too:
```
$\begin{pmatrix}  
1 & 2 & 3 \\ 
4 & 5 & 6
\end{pmatrix}$
```
$\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6
\end{pmatrix}$ 

Brackets:  
```
$\begin{bmatrix}  
1 & 2 \\ 
4 & 5   
\end{bmatrix}$  
```
$\begin{bmatrix}
1 & 2 \\
4 & 5 
\end{bmatrix}$  

Very small:
```
$\left[\begin{smallmatrix}  
1 & 2 \\
4 & 5   
\end{smallmatrix}\right]$  
```
$\left[\begin{smallmatrix}
1 & 2 \\
4 & 5 
\end{smallmatrix}\right]$





[^1]: You may have noticed the footnote above. Here it is. Some of these examples were taken from [here](https://rstudio-pubs-static.s3.amazonaws.com/256225_63ebef4029dd40ef8e3679f6cf200a5a.html#).



